---
title: "Practical Machine Learning - Project"
author: "Joao Rei"
date: "Sunday, March 20, 2015"
output: html_document
---

# Executive Summary

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. 
In this project, the goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.The objective is to predict the manner in which they did the exercise. This is the "classe" variable in the training set.


# Getting and Cleaning Data

The training data for this project is available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv
The test data is available here: 
https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

Both were manually downloaded and placed in the working directory.


```{r echo=FALSE}
setwd("C:/DISCO_D/ORGANIZADO/Coursera/__Coursera/08_Data Science Specialization/08_Practical Machine Learning/PROJECT")

```

The training and testing data were then loaded to R. The first 7 columns of the datasets were removed because they dont have any prediction value.
Any column with NAs was also removed.
The training dataset was also split to build a verification dataset (30% of data from training).


```{r}
library(caret)

training = read.csv("pml-training.csv", na.strings=c("NA",""))
testing = read.csv("pml-testing.csv",na.strings=c("NA",""))

training = training[,8:length(training)]
testing = testing[,8:length(testing)]

training_NAs <- apply(training, 2, function(x) {sum(is.na(x))})
testing_NAs <- apply(testing, 2, function(x) {sum(is.na(x))})

training = training[,which(training_NAs == 0)]
testing = testing[,which(training_NAs == 0)]

inTrain <- createDataPartition(training$classe, p=0.70, list=FALSE)
training <- training[inTrain,]
validation <- training[-inTrain,]

```


# Model

For this prediction a RandomForest model was choosen. Caret library was used to fit the model. Accuracy was used to select the optimal model using  the largest value.
The final value used for the model was mtry = 27.
This Accuracy was calculated using Resampling Cross-Validated.


```{r cache=TRUE}

traincontrol <- trainControl(method = "cv", number = 5)
model = train(classe ~ . ,data=training, method="rf",trControl=traincontrol)

```

# Out of Sample Error

Accuracy of 0.9914102 or a small 0.86% out of sample error was calculated from 
Resampling Cross-Validated, (mtry=27).
Out of Sample from validation data was also calculated, with Accuracy = 0.9973 or 0.37% out of sample error.


```{r}
model
pred = predict(model, validation)
confusionMatrix(pred, validation$classe)

```


# Prediction

The rf model was then used to predict 20 test cases.
The testing dataset had a feature problem_id that was removed before running predict function.

```{r}

testing2 = testing
testing2$problem_id = NULL
testing2$classe = rep("A",20)

test_prediction = predict(model, testing2)

str(test_prediction)


pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}
pml_write_files(test_prediction)


```


